\large How resource abundance and stochasticity affect organisms' space-use requirements

### 2023-09-20 writing/manuscript.Rmd
updated based on Chrirs's comments

\noindent Simon Wood: simon.wood@bath.edu

\noindent Chris H. Fleming: chris.h.fleming@gmail.com

@@

@rizzuto_forage_2021 found that the space-use requirements of snowshoe hares (*Lepus americanus*) increased with average carbon to nitrogen ratio (C:N, a measure of N scarcity) in lowbush blueberry as well as the coefficient of variation (the standard deviation over the mean, $\frac{\sqrt{\sigma^2(t)}}{\mu(t)}$) in C:N. However, it is hard to determine how $\sigma^2(t)$ affected the hares' space-use requirements since the coefficient of variation is a function of both $\mu(t)$ and $\sigma^2(t)$, so the effects of the two variables are confounded. Similarly, @mueller_how_2011 suggested that ecosystem variability (measured as spatial semivariance in NDVI) causes ungulates to move more and adopt more nomadic behaviors, but it is hard to determine whether the higher semivariances in NDVI are due to an increase in $\var(R)$ or simply spatial changes in $\e(R)$ (e.g., increased seasonality or continentality). Likewise, @herfindal_prey_2005 found that the home ranges of Canadian lynx (*Lynx canadensis*) decreased with seasonality, but this may be an artifact of a lower $\mu(t)$ in highly seasonal environments, rather than stochasticity in $R$. A subsequent analysis by @nilsen_can_2005 of the lynx data of @herfindal_prey_2005 along with data on wolverines (*Gulo gulo*), bobcats (*Lynx rufus*), and coyotes (*Canis latrans*) found that carnivores respond differently to seasonality, which suggests that different species may have different perceptions of the changes in $R$ over the seasons. Collectively, these studies provide support towards the effects of both $\e(R)$ and $\var(R)$ on organisms' space-use requirements, but the lack of a unifying framework across these studies makes it challenging to synthesize their findings. Thus, it is important to disentangle the effects of changes in $\e(R)$, such as seasonal trends in $\mu(t)$, and stochasticity ($\var(R)$). The need to account for changes in both $\e(R)$ and $\var(R)$ is compounded by recent changes in climate, which expose species to novel situations and increasingly common and extreme stochastic events [@intergovernmental_panel_on_climate_change_climate_2023; @noonan_situ_2018; @yao_emergence_2022].

>>>>

\noindent Chris H. Fleming: christen.fleming@ucf.edu

@@

@rizzuto_forage_2021 found that the space-use requirements of snowshoe hares (*Lepus americanus*) increased with average carbon to nitrogen ratio (C:N, a measure of N scarcity) in lowbush blueberry as well as the coefficient of variation (the standard deviation divided by the mean, $\frac{\sqrt{\sigma^2(t)}}{\mu(t)}$) in C:N. However, it is hard to determine how $\sigma^2(t)$ affected the hares' space-use requirements since the coefficient of variation is a function of both $\mu(t)$ and $\sigma^2(t)$, so the effects of the two variables are confounded. Similarly, @mueller_how_2011 suggested that ecosystem variability (measured as spatial semivariance in NDVI) causes ungulates to move more and adopt more nomadic behaviors, but it is hard to determine whether the higher semivariances in NDVI are due to an increase in $\var(R)$ or simply spatial changes in $\e(R)$ (e.g., increased seasonality or continentality). Likewise, @herfindal_prey_2005 found that the home ranges of Canadian lynx (*Lynx canadensis*) decreased with seasonality, but this may be an artifact of a lower $\mu(t)$ in highly seasonal environments, rather than stochasticity in $R$. A subsequent analysis by @nilsen_can_2005 of the lynx data of @herfindal_prey_2005 along with data on wolverines (*Gulo gulo*), bobcats (*Lynx rufus*), and coyotes (*Canis latrans*) found that carnivores respond differently to seasonality, which suggests that different species may have different perceptions of the changes in $R$ over the seasons. Collectively, these studies provide support towards the effects of both $\e(R)$ and $\var(R)$ on organisms' space-use requirements, but the lack of a unifying framework across these studies makes it challenging to synthesize their findings. Thus, it is important to disentangle the effects of changes in $\e(R)$, such as seasonal trends in $\mu(t)$, and stochasticity ($\var(R)$). The need to account for changes in both $\e(R)$ and $\var(R)$ is compounded by recent changes in climate, which expose species to novel situations and increasingly common and extreme stochastic events [@intergovernmental_panel_on_climate_change_climate_2023; @noonan_situ_2018; @yao_emergence_2022].
